\section{Introduction}\label{sec:introduction}

\begin{comment}
1. Changed "composers, music theorists and musicologists" to just "composers". How do we feel about this?
\end{comment}

Abjad\footnote{www.projectabjad.org} is an interactive software system designed
to help composers build up complex pieces of music notation in an iterative and
incremental way. Abjad is implemented in the Python\footnote{www.python.org}
programming language and architected as an object-oriented collection of
packages, classes and functions. Users can visualize their work as
publication-quality score at all stages of the compositional process using
Abjad's interface to the LilyPond\footnote{www.lilypond.org} music notation
package. Abjad is open-source software available for free download from the
Python Package Index.\footnote{pypi.python.org}

While many environments for both notation and sound production have arisen
within the last twenty years, the following discussion focuses solely on the
production of notation: Abjad enables composers to express both low- and
high-level compositional ideas by extending a widely used programming language
to provide a sufficiently detailed object model of common practice musical
notation. To minimize the restriction of artistic thought's infinite
possibility while maximizing the ability to specify elegantly any arbitrary
symbolic relationship, Abjad does this without prescribing explicit or implicit
models of music or composition: Abjad defines composition narrowly as the act
of creating a document via the encoded aggregation of notational symbols.

\begin{comment}
The current version of Abjad implements 491 public classes and 324 public
functions.
\end{comment}